Exploring the conformational dynamics of a large molecule in solution using an external electric field: A Nonequilibrium Molecular Dynamics Simulation

### 2013-08-27 writedown/npt-neq/alanine.tex
Luigi's revise to the conclusion

The method we employ is the dynamical nonequilibrium molecular dynamics simulation (D-NEMD).
The method was firstly developed by one of us (GC) and coworkers~\cite{ciccotti1975direct, ciccotti1979thought}.
It has been recently applied, with success, to the study of
hydrodynamics~\cite{orlandini2011hydrodynamics,
  orlandini2011hydrodynamics-01}. The essential feature of D-NEMD is the following: It provides a way of evaluating
nonequilibrium observables by using a properly determined ensemble of nonequilibrium MD
simulations.  In this context, the method can be used for analyzing the
conformational changes of a solvated molecule, subject to a time-dependent EF.
In the present paper, we study a small peptide, alanine dipeptide, as a model for showing how the method can be used. 
The choice of a small peptide, rather than a larger one, is done firstly because it
is easier to draw clear conclusions on the nonequilibrium properties
of a single peptide by excluding a possible interplay between
different peptide segments along a chain.
Secondly, the computational cost required is rather massive and thus the testing of the various methodolgy developments proposed here would not be easy to accomplish for large molecules. Of course, given a positive coming out for such a model system, the aim is that of approaching systems of growing complexity.
The paper is organized as follows:
We firstly describe the idea of D-NEMD, then
introduce the physical
observables for analyzing the nonequilibrium properties. The numerical results of alanine
dipeptide under static and sine-shaped oscillatory EF are
finally presented. By calculating the nonequilibrium properties from D-NEMD
simulations, we are able to address the question of how the peptide
changes the conformation in a EF (conformational space), when do these changes happens (time scale),
and how likely they are (probability). 
To our knowledge, this study is one of the few (if not the first)  where non equilibrium is explicitly considered for a large molecule in solution bubject to time-dependent EF.

\section{Methodology}
\subsection{Nonequilibrium molecular dynamics simulation}
In this section we briefly review the approach for performing dynamical nonequilibrium
molecular dynamics (D-NEMD) \cite{ciccotti1975direct, ciccotti1979thought,
  orlandini2011hydrodynamics, orlandini2011hydrodynamics-01}.
In the following we denote the macroscopic observable by $O(t)$. If at $t$
the configurational probability distribution is $\rho(\vect x, t)$, where
$\vect x$ is the phase space variable, then the observable can be
expressed by
\begin{align}\label{eqn:tmp1}
  O(t) = \int d\vect x\, \hat O(\vect x)\rho(\vect x, t)  = \langle \hat O(\vect x), \rho(\vect x, t)\rangle,
\end{align}
we refer to $\hat O (\vect x)$ as the microscopic observable, 
which is measured at each point in the phase space $\vect x$.
Here we always assume that the initial probability distribution
$\rho(\vect x, 0)$ is known and is identical to the equilibrium distribution of the system
without EF.
The bracket on the right hand side of Eq.~\eqref{eqn:tmp1} denotes, as usual, the inner product in the
phase space.  We assume tthat he dynamics of the system is governed by the
Hamiltonian equation, i.e. $\dot {\vect x} = J \cdot \nabla_{\vect x}
\mathcal H(\vect x)$, where $\mathcal H$ is the Hamiltonian, and $J$ is
the symplectic matrix. The Liouville equation for the probability
distribution writes:
\begin{align}\label{eqn:tmp2}
  \frac{\partial \rho(\vect x, t)}{\partial t} = - iL(t) \rho(\vect x, t),
\end{align}
where $iL(t) = \{\cdot, \mathcal H\}$ is the Liouville operator.
Eq.~\eqref{eqn:tmp2}
can be formally solved by $\rho(\vect x, t) = e^{-iL(t)} \rho(\vect x, 0)$.
On the other hand
\begin{align}
  \frac{d \hat O(\vect x(t))}{dt} = \nabla_{\vect x}\hat O\cdot \dot{\vect x}
  = \nabla_{\vect x}\hat O\cdot J\cdot \nabla_{\vect x}\mathcal H
  = iL(t) \hat O (\vect x(t))
\end{align}
This equation can be formally solved by $\hat O(\vect x(t)) = e^{iL(t)} \hat O(\vect x(0))$, therefore,
\begin{align}\nonumber
  O(t) & = \langle \hat O(\vect x), \rho(\vect x, t)\rangle
  = \langle \hat O(\vect x), e^{-iL(t)} \rho(\vect x, 0)\rangle
  = \langle e^{iL(t)}\hat O(\vect x), \rho(\vect x, 0)\rangle\\\label{eqn:tmp4}
  &= \langle \hat O(\vect x(t)), \rho(\vect x, 0)\rangle
\end{align}
Since we assume that the system starts from the distribution of equilibrium (without EF)
 Eq.~\eqref{eqn:tmp4} expresses the fact that the 
observable $O(t)$ calculated in a situation of nonequilibrium is equal to the ensemble average of
the microscopic observable measured along trajectories, the
initial configurations of which are subject to the \emph{equilibrium}
distribution. In practical terms we proceed by first running an equilibrium MD simulation
to generate a sample of configurations. Next we employ these configurations as
initial configurations and, for each, the Hamiltonian dynamics is integrated until
time $t$. These trajectories are also called \emph{branching
  trajectories} in this paper (see Fig.\ref{branching} for a pictorial representation). 
Finally, the value of the macroscopic observable at time $t$ is
calculated by averaging the microscopic observable measured at each time $t$ along each branching trajectory.

@@

% The aforementioned D-NEMD simulation algorithm
% does not describe how the boundary of the system should be set up.
% The very naive idea of doing the simulation
% is to simulate a infinitely large isolated system, or any large enough
% isolated subsystem of it.
% In practice, 
% Therefore, we are interested in a 
For this study We are interested in 
nonthermal~\cite{delaHoz2005microwaves} effects of the EF
on the configuration of the solvated alanine dipeptide.
For this reason a proper control of the temperature in a nonequilibrium system
becomes an important issue; in fact it was argued that  an
accurate control of the internal reaction temperature is essential for being able to
perform reproducible experiments~\cite{damm2012can}.
A straigthforward (and realistic) implementation of the temperature control consists in 
embedding the alanine into an infinitely large solvent environment, and 
the EF is only applied to the neighborhood of the alanine. In this case
any extra heat generated by the EF can be effectively
dispersed in the large solvent environment.
In this way one can identify the effect of the EF on the conformational space of the solvated molecule,
without the artifact that the surrounding solvant cannot release the additional energy of the EF.
However, in practice, in most of the cases the computational cost of having large solvant environment is rather massive if not prohibitive.
Therefore, it is of interest to substitute the infinitely large
solvent environment with a smaller affordable system and at the same time accurately reproduce the nonequilibrium
physical properties of interest.

As it is well known, in MD, for equilibrium situation, the solution is to couple
the system to a an external thermostat, under which the desired ensemble
(canonical in this case) can be
sampled by simulating a finite size system with periodic boundary
conditions.
However, coupling the system to a thermostat for non equilibrium situations is a delicate issue.
In fact, since a thermostat is explicitly designed to bring the system at equilibrium, any perturbation produced by the EF, would be mostly adsorbed by the thermostat. This means that the physics of the original Hamiltonian dynamics with the EF would be lost and the response to the perturbation becomes unphysical. 
%will disturb the original Hamiltonian dynamics. This is
%in general problematic, 
%because the dynamics  under thermostating is artificial designed,
%so that the equilibrium distribution can be correctly produce, but not
%the nonequilibrium properties of system.
% dynamics of the system is no longer Hamiltonian, and is
% artificial designed to reach the equilibrium probability distribution.
% However, for a nonequilibrium system, the \emph{dynamics} properties
% will be distrubed by the thermostat, because the
% dynamics of the system is no longer Hamiltonian, and is
% artificial designed to reach the equilibrium probability distribution.
% For a nonequilibrium simulation, simply couple the system
% to a thermostat may not be a good idea, because the \emph{dynamics}
% generated by the thermostat is artificial, and deviation from the
% Hamiltonian dynamics under investigation will introduce artificial
% a effects in the nonequilibrium observations.
In this section we propose a procedure to solve the problem: we
observe the Hamiltonian dynamics (without thermostat) of the alanine 
itself and the water molecules in the closer solvation shells is
crucial in order to determine the response of the observable of interest
under the perturbation of the EF (in this work, it is the conformational change of alanine).
Instead the detailed dynamics of
the water molecules far away is not of primary importance, that is the region beyond the first solvation shells plays the role of a thermodynamic bath only. 
Therefore, instead of simulating an infinitely
large system, we only simulate a finite size system with periodic
boundary condition with the following characteristics. The simulating region is divided in two
subregions: a spherically defined region around the alanine dipeptide, centered at the
alpha-carbon of radius $R_{ex}$. Here the dynamics is
Hamiltonian and molecules are not subject to a thermosat; we refer to such subregion as \emph{the Hamiltonian dynamics
region}. Instead beyond $R_{ex}$, the dynamics of water is coupled
to a Langevin thermostat, and this region is called \emph{the thermostated
region} (see Fig.~\ref{fig:tmp2}). For technical convenience we
fix the position of the alpha-carbon in space, so that the alanine is always
located at the center of the simulation region.
Therefore, a realistic non equilibrium  dynamics is preseved in the Hamiltonian dynamics region (where the properties of interest are observed)
and possible direct artificial effects of the thermostating process in the region outside are negligible because of the finite correlation
range of liquid water.
At the same time, the thermostated region works as a infinitely large
environment that effectively absorbs the extra heat produced by the EF.
The validity of the above
statement will be later on carefully checked by a series of numerical tests where it is shown that
the response of observables in nonequilibrium (i.e. under the effect of the EF) does not depend on the size of the system
and on the size of the Hamiltonian dynamics subregion, provided that they are
reasonably large.

@@

From Fig.~\ref{fig:tmp4}, it is obvious that the population in each
range of conformations changes when the external EF is applied.
For instance, the population of conformation $C$ vanishes in equilibrium,
but plays an important role when the EF is applied. We
therefore want to study how the alanine changes from one conformation
to the other, how fast is the change, and the relations to the 
the EF. 
Firstl, we calculate the probability of being in a certain conformation range,
defined as:
\begin{align}
  P_I(t) = \mathbb P (\vect x_t \in I), \quad  I \in \{A_1, A_2, B_1, B_2, C\},
\end{align}
Notice that the probability is time dependent, being an observable in nonequilibrium situation.
% which is the time dependent probability of the alanine being in metastable
% state $I$.
% Secondly, we study the
% forward and backward transition probability of the metastable conformations, which is defined by
% \begin{align}
%   P^{\fwd}_{J,I}(t) & = \mathbb P( \vect x_{t+\Delta t} \in J | \vect x_t \in I) \\
%   P^{\bwd}_{J,I}(t) & = \mathbb P( \vect x_{t-\Delta t} \in J | \vect x_t \in I) 
%   % \qquad I, J \in \{A_1, A_2, B_1, B_2, C\},
% \end{align}
% where $\Delta t$ is the time-scale of the observation.
% For a revserible and homogeneous Markov process, it is easy to show
% that $P^{\fwd}_{J,I} = P^{\bwd}_{J,I}$, and both of them are time independent. For a general nonequilibrium
% process considered by the present work, this relation is genarlly not guarenteed.
Second,
we study the net probability flux from conformation $I$ to $J$, defined by:
\begin{align}\nonumber
  F_{J,I}(t) & =\lim_{\Delta t\rightarrow 0} \frac1{\Delta t} [\mathbb P( \vect x_{t-\Delta t} \in J, \vect x_t \in I) - \mathbb P( \vect x_{t-\Delta t} \in I, \vect x_t \in J)], \\\label{eqn:tmp7}
  & J,I \in \{A_1, A_2, B_1, B_2, C\}.
\end{align}
$F_{J,I}(t)$ is essentially the rate of how much conformation $J$ changes
into $I$ at time $t$. A positive value indicates the
net flux from $J$ to $I$,
while a negative value indicates a net flux from $I$ to $J$.
Through $F_{J,I}(t)$, the original overall nonequilibrium process $\vect x_t$
is reduced to the probabilistic link between discretized conformations $\{A_1, A_2, B_1, B_2, C\}$.
If the resulting reduced process 
is time-reversible, then it is obvious that
$ F_{J,I}(t) = 0$. In an other word, the quantity $ F_{J,I}(t) $ represents
the irreversibility of the reduced process on
$\{A_1, A_2, B_1, B_2, C\}$.
In practice, to measure $ F_{J,I}(t)$, one need first to approximate the limit
in Eq.~\eqref{eqn:tmp7} by the finite difference,
then calculate the joint probabilities on
the right-hand-side from the molecular trajectories.
In this context, for a precise finite difference approximation a smaller $\Delta t$ is
preferable although the statistical uncertainty of measuring the joint probabilities
is higher. In practice,
we use a $\Delta t$ that is small enough for a precise finite difference approximation,
and not small enough so that the joint probabilities can be calculated with acceptable accuracy.
In the case of oscillatory EF, the net probability flux is also highly oscillatory,
thus it is convenient to study the integrated probability flux:
\begin{align}
  Q_{J,I} (t) = \int_0^t F_{J,I}(\tau)d \tau,
\end{align}
this expresses the average over time of the effects of the EF on the conformation of the molecule.

@@

\begin{table}
  \centering
  \begin{tabular*}{0.4\textwidth}{@{\extracolsep{\fill}}lc}\hline\hline
    Direction        & time scale [ ps ] \\\hline
    $A_1\rightarrow A_2$        &       $\sim 10$      \\
    $B_1\rightarrow B_2$        &       $\sim 10$      \\    
    $B_2\rightarrow A_1,\ A_2,\ C$        &       $\sim 50$      \\    
    $A_2\rightarrow A_1$        &       $\sim 100$      \\
    $B_2\rightarrow B_1$        &       $\sim 100$      \\    
    $B_1\rightarrow A_1,\ C$        &       $\sim 500$      \\    \hline\hline
  \end{tabular*}
  \caption{A list of main probabilty fluxes and the corresponding time scales observed in the constant EF case.}
  \label{tab:tmp1}
\end{table}

@@

The Fig.~\ref{fig:tmp3}
shows that the $x$-component of the dipole moment of the alanine dipeptide
reaches 85\% of the maximum value in only 20~ps, which is comparable to
the warm-up time $t_{warm}$, and is 25 times
smaller than the slowest time scale of the conformational relaxation.
Then in the following 400~ps, the dipole slowly relaxes to the 
maximum value, that is, 
6.8~Debye, and  corresponds to a dipole energy of ca. $-13.6$~kJ/mol.
We calculate the averaged dipole moment of different conformations
under the constant EF, and they are 6.8, 6.0, 5.1, 3.1 and 7.1~Debye
for $A_1$, $A_2$, $B_1$, $B_2$ and $C$, respectively.
The dipole energy difference between 3.1 and 7.1~Debye is roughly 8.0~kJ/mol.
The free energy barriers between the
conformations were reported to be of order $20\sim 40$~kJ/mol~\cite{bohner2012algorithm},
which is comparable to the dipole energy difference. 
Therefore, under a constant EF, the system will be driven towards those
conformations with higher dipole moment, because the
energy of the system will be lowered.  This is the reason why
we observe that the $\beta$-sheet conformations with lower dipole moment
are driven towars $\alpha$-helix conformations, 
and $A_1$ and $C$ are conformations with higher probability, because their dipole moments are the highest
among all conformations. Moreover, the $B_2$ conformation vanishes because
its dipole is significantly lower than other conformations.
% This also explains the two-time scale behavior of the molecular dipole
% moment.
From 0 to 20~ps, the probability of
lowest dipole conformation $B_2$ quickly
vanishes. We observe a sharp increment of the average molecular dipole and finally, approaching  500~ps, the conformation slowly migrates from
$B_1$ to $A_1$ and $C$. Moreover, since the dipole moment $A_1$ and $C$
is higher than $B_1$, the average molecular dipole moment slowly increases
to the nonequilibrium steady value.
\subsection{Response of the orientational order parameter}

We also consider the orientation of the alanine dipeptide in space (w.r.t. the direction of EF) as a function
of time.
The geometric direction of the alanine is defined by the red vector in
Fig.~\ref{fig:tmp3}, which is the angle bisector of the two black vectors.
The black vectors are defined in such a way that they connect the $\alpha$-carbon and the carbons
on the methyl groups. We define the orientational order parameter by
\begin{align}
  S_\theta = \langle 3\sin^2\theta - 2\rangle,
\end{align}
where $\theta$ is the angle between the orientation of the alanine dipeptide
and the direction of the EF. The order parameter
indicate how perpendicular does the molecule lie w.r.t. the EF.
If the molecule is perfectly perpendicular to the EF, then $S_\theta = 1$;
If the molecule has no directional
preference , then $S_\theta = 0$;
If the molecule lies perfectly parallel to the EF, then $S_\theta = -2$.
From $t=0$ to roughly 10~ps, the orientational order parameter fastly
decreases from 0 to $-0.15$, which means a weak alignment of the molecule to the external field.
Then from $t=0$ to 500~ps, the molecule slowly changes to the orientation
that is perpendicular to the EF.
This orientational change is linked to the observed slowest
time scale of the molecular conformational change.

@@

Just as described in Sec.~\ref{sec:tmp2b}, we perform nonequilibrium MD
simulations in a finite size periodic system, and further divided it
into a Hamiltonian dynamics region and a thermostated region.
Since the periodic boundary conditions and the division of the system
are technical (not physical) approximations to the real system with dynamics-preserveing thermal control, 
the effect of the finite-size in 
the system settings should be carefully checked.
Therefore, we
perform two additional simulations: one with box dimension $L=4.0$~nm (box size on all directions are the same) and a
Hamiltonian dynamics region of radius $R_{ex} = 1.0$~nm. A second simulation is done with box size $L=4.0$~nm and a Hamiltonian dynamics region of
radius $R_{ex} = 1.5$~nm. We compare the results of these two simulations to those of the system we have used so far, i.e. $L=2.7$~nm and $R_{ex} = 1.0$~nm.
Fig.~\ref{fig:tmp7} presents a satisfactory consistency in
the simulation results of the three systems, which implies that the finite-size effects are
actually negligible.

>>>>

The method we employ is the dynamical nonequilibrium molecular dynamics simulation (D-NEMD).
The method was firstly developed by one of us (GC) and coworkers~\cite{ciccotti1975direct, ciccotti1979thought}.
It has been recently applied, with success, to the study of
hydrodynamics~\cite{orlandini2011hydrodynamics,
  orlandini2011hydrodynamics-01}. The essential feature of D-NEMD is the following: It provides a way of evaluating
nonequilibrium observables by using a properly determined ensemble of nonequilibrium MD
simulations.  In this context, the method can be used for analyzing the
conformational changes of a solvated molecule, subject to a time-dependent EF.
In the present paper, we study a small peptide, alanine dipeptide, as a model for showing how the method can be used. 
The choice of a small peptide, rather than a larger one, is done firstly because it
is easier to draw clear conclusions on the nonequilibrium properties
of a single peptide by excluding a possible interplay between
different peptide segments along a chain.
Secondly, the computational cost required is rather massive and thus the testing of the various methodology developments proposed here would not be easy to accomplish for large molecules. Of course, given a positive coming out for such a model system, the aim is that of approaching systems of growing complexity.
The paper is organized as follows:
We firstly describe the idea of D-NEMD, then
introduce the physical
observables for analyzing the nonequilibrium properties. The numerical results of alanine
dipeptide under static and sine-shaped oscillatory EF are
finally presented. By calculating the nonequilibrium properties from D-NEMD
simulations, we are able to address the question of how the peptide
changes the conformation in a EF (conformational space), when do these changes happens (time scale),
and how likely they are (probability). 
To our knowledge, this study is one of the few (if not the first)  where non equilibrium is explicitly considered for a large molecule in solution subject to time-dependent EF.

\section{Methodology}
\subsection{Nonequilibrium molecular dynamics simulation}
In this section we briefly review the approach for performing dynamical nonequilibrium
molecular dynamics (D-NEMD) \cite{ciccotti1975direct, ciccotti1979thought,
  orlandini2011hydrodynamics, orlandini2011hydrodynamics-01}.
In the following we denote the macroscopic observable by $O(t)$. If at $t$
the configurational probability distribution is $\rho(\vect x, t)$, where
$\vect x$ is the phase space variable, then the observable can be
expressed by
\begin{align}\label{eqn:tmp1}
  O(t) = \int d\vect x\, \hat O(\vect x)\rho(\vect x, t)  = \langle \hat O(\vect x), \rho(\vect x, t)\rangle,
\end{align}
we refer to $\hat O (\vect x)$ as the microscopic observable, 
which is measured at each point in the phase space $\vect x$.
Here we always assume that the initial probability distribution
$\rho(\vect x, 0)$ is known and is identical to the equilibrium distribution of the system
without EF.
The bracket on the right hand side of Eq.~\eqref{eqn:tmp1} denotes, as usual, the inner product in the
phase space.  We assume that he dynamics of the system is governed by the
Hamiltonian equation, i.e. $\dot {\vect x} = J \cdot \nabla_{\vect x}
\mathcal H(\vect x)$, where $\mathcal H$ is the Hamiltonian, and $J$ is
the symplectic matrix. The Liouville equation for the probability
distribution writes:
\begin{align}\label{eqn:tmp2}
  \frac{\partial \rho(\vect x, t)}{\partial t} = - iL(t) \rho(\vect x, t),
\end{align}
where $iL(t) = \{\cdot, \mathcal H\}$ is the Liouville operator.
Eq.~\eqref{eqn:tmp2}
can be formally solved by $\rho(\vect x, t) = e^{-iL(t)} \rho(\vect x, 0)$.
On the other hand
\begin{align}
  \frac{d \hat O(\vect x(t))}{dt} = \nabla_{\vect x}\hat O\cdot \dot{\vect x}
  = \nabla_{\vect x}\hat O\cdot J\cdot \nabla_{\vect x}\mathcal H
  = iL(t) \hat O (\vect x(t))
\end{align}
This equation can be formally solved by $\hat O(\vect x(t)) = e^{iL(t)} \hat O(\vect x(0))$, therefore,
\begin{align}\nonumber
  O(t) & = \langle \hat O(\vect x), \rho(\vect x, t)\rangle
  = \langle \hat O(\vect x), e^{-iL(t)} \rho(\vect x, 0)\rangle
  = \langle e^{iL(t)}\hat O(\vect x), \rho(\vect x, 0)\rangle\\\label{eqn:tmp4}
  &= \langle \hat O(\vect x(t)), \rho(\vect x, 0)\rangle
\end{align}
Since we assume that the system starts from the distribution of equilibrium (without EF)
 Eq.~\eqref{eqn:tmp4} expresses the fact that the 
observable $O(t)$ calculated in a situation of nonequilibrium is equal to the ensemble average of
the microscopic observable measured along trajectories, the
initial configurations of which are subject to the \emph{equilibrium}
distribution. In practical terms we proceed by first running an equilibrium MD simulation
to generate a sample of configurations. Next we employ these configurations as
initial configurations and, for each, the Hamiltonian dynamics is integrated until
time $t$. These trajectories are also called \emph{branching
  trajectories} in this paper (see Fig.\ref{branching} for a pictorial representation). 
Finally, the value of the macroscopic observable at time $t$ is
calculated by averaging the microscopic observable measured at each time $t$ along each branching trajectory.

@@

% The aforementioned D-NEMD simulation algorithm
% does not describe how the boundary of the system should be set up.
% The very naive idea of doing the simulation
% is to simulate a infinitely large isolated system, or any large enough
% isolated subsystem of it.
% In practice, 
% Therefore, we are interested in a 
For this study We are interested in 
nonthermal~\cite{delaHoz2005microwaves} effects of the EF
on the configuration of the solvated alanine dipeptide.
For this reason a proper control of the temperature in a nonequilibrium system
becomes an important issue; in fact it was argued that  an
accurate control of the internal reaction temperature is essential for being able to
perform reproducible experiments~\cite{damm2012can}.
A straightforward (and realistic) implementation of the temperature control consists in 
embedding the alanine into an infinitely large solvent environment, and 
the EF is only applied to the neighborhood of the alanine. In this case
any extra heat generated by the EF can be effectively
dispersed in the large solvent environment.
In this way one can identify the effect of the EF on the conformational space of the solvated molecule,
without the artifact that the surrounding solvent cannot release the additional energy of the EF.
However, in practice, in most of the cases the computational cost of having large solvent environment is rather massive if not prohibitive.
Therefore, it is of interest to substitute the infinitely large
solvent environment with a smaller affordable system and at the same time accurately reproduce the nonequilibrium
physical properties of interest.

As it is well known, in MD, for equilibrium situation, the solution is to couple
the system to a an external thermostat, under which the desired ensemble
(canonical in this case) can be
sampled by simulating a finite size system with periodic boundary
conditions.
However, coupling the system to a thermostat for non equilibrium situations is a delicate issue.
In fact, since a thermostat is explicitly designed to bring the system at equilibrium, any perturbation produced by the EF, would be mostly adsorbed by the thermostat. This means that the physics of the original Hamiltonian dynamics with the EF would be lost and the response to the perturbation becomes unphysical. 
%will disturb the original Hamiltonian dynamics. This is
%in general problematic, 
%because the dynamics  under thermostating is artificial designed,
%so that the equilibrium distribution can be correctly produce, but not
%the nonequilibrium properties of system.
% dynamics of the system is no longer Hamiltonian, and is
% artificial designed to reach the equilibrium probability distribution.
% However, for a nonequilibrium system, the \emph{dynamics} properties
% will be distrubed by the thermostat, because the
% dynamics of the system is no longer Hamiltonian, and is
% artificial designed to reach the equilibrium probability distribution.
% For a nonequilibrium simulation, simply couple the system
% to a thermostat may not be a good idea, because the \emph{dynamics}
% generated by the thermostat is artificial, and deviation from the
% Hamiltonian dynamics under investigation will introduce artificial
% a effects in the nonequilibrium observations.
In this section we propose a procedure to solve the problem: we
observe the Hamiltonian dynamics (without thermostat) of the alanine 
itself and the water molecules in the closer solvation shells is
crucial in order to determine the response of the observable of interest
under the perturbation of the EF (in this work, it is the conformational change of alanine).
Instead the detailed dynamics of
the water molecules far away is not of primary importance, that is the region beyond the first solvation shells plays the role of a thermodynamic bath only. 
Therefore, instead of simulating an infinitely
large system, we only simulate a finite size system with periodic
boundary condition with the following characteristics. The simulating region is divided in two
subregions: a spherically defined region around the alanine dipeptide, centered at the
alpha-carbon of radius $R_{ex}$. Here the dynamics is
Hamiltonian and molecules are not subject to a thermostat; we refer to such subregion as \emph{the Hamiltonian dynamics
region}. Instead beyond $R_{ex}$, the dynamics of water is coupled
to a Langevin thermostat, and this region is called \emph{the thermostated
region} (see Fig.~\ref{fig:tmp2}). For technical convenience we
fix the position of the alpha-carbon in space, so that the alanine is always
located at the center of the simulation region.
Therefore, a realistic non equilibrium  dynamics is preserved in the Hamiltonian dynamics region (where the properties of interest are observed)
and possible direct artificial effects of the thermostating process in the region outside are negligible because of the finite correlation
range of liquid water.
At the same time, the thermostated region works as a infinitely large
environment that effectively absorbs the extra heat produced by the EF.
The validity of the above
statement will be later on carefully checked by a series of numerical tests where it is shown that
the response of observables in nonequilibrium (i.e. under the effect of the EF) does not depend on the size of the system
and on the size of the Hamiltonian dynamics subregion, provided that they are
reasonably large.

@@

From Fig.~\ref{fig:tmp4}, it is obvious that the population in each
range of conformations changes when the external EF is applied.
For instance, the population of conformation $C$ vanishes in equilibrium,
but plays an important role when the EF is applied. We
therefore want to study how the alanine changes from one conformation
to the other, how fast is the change, and the relations to the 
the EF. 
First, we calculate the probability of being in a certain conformation range,
defined as:
\begin{align}
  P_I(t) = \mathbb P (\vect x_t \in I), \quad  I \in \{A_1, A_2, B_1, B_2, C\},
\end{align}
Notice that the probability is time dependent, being an observable in nonequilibrium situation.
% which is the time dependent probability of the alanine being in metastable
% state $I$.
% Secondly, we study the
% forward and backward transition probability of the metastable conformations, which is defined by
% \begin{align}
%   P^{\fwd}_{J,I}(t) & = \mathbb P( \vect x_{t+\Delta t} \in J | \vect x_t \in I) \\
%   P^{\bwd}_{J,I}(t) & = \mathbb P( \vect x_{t-\Delta t} \in J | \vect x_t \in I) 
%   % \qquad I, J \in \{A_1, A_2, B_1, B_2, C\},
% \end{align}
% where $\Delta t$ is the time-scale of the observation.
% For a revserible and homogeneous Markov process, it is easy to show
% that $P^{\fwd}_{J,I} = P^{\bwd}_{J,I}$, and both of them are time independent. For a general nonequilibrium
% process considered by the present work, this relation is genarlly not guarenteed.
Second,
we study the net probability flux from conformation $I$ to $J$, defined by:
\begin{align}\nonumber
  F_{J,I}(t) & =\lim_{\Delta t\rightarrow 0} \frac1{\Delta t} [\mathbb P( \vect x_{t-\Delta t} \in J, \vect x_t \in I) - \mathbb P( \vect x_{t-\Delta t} \in I, \vect x_t \in J)], \\\label{eqn:tmp7}
  & J,I \in \{A_1, A_2, B_1, B_2, C\}.
\end{align}
$F_{J,I}(t)$ is essentially the rate of how much conformation $J$ changes
into $I$ at time $t$. A positive value indicates the
net flux from $J$ to $I$,
while a negative value indicates a net flux from $I$ to $J$.
Through $F_{J,I}(t)$, the original overall nonequilibrium process $\vect x_t$
is reduced to the probabilistic link between discretized conformations $\{A_1, A_2, B_1, B_2, C\}$.
If the resulting reduced process 
is time-reversible, then it is obvious that
$ F_{J,I}(t) = 0$. In an other word, the quantity $ F_{J,I}(t) $ represents
the irreversibility of the reduced process on
$\{A_1, A_2, B_1, B_2, C\}$.
In practice, to measure $ F_{J,I}(t)$, one need first to approximate the limit
in Eq.~\eqref{eqn:tmp7} by the finite difference,
then calculate the joint probabilities on
the right-hand-side from the molecular trajectories.
In this context, for a precise finite difference approximation a smaller $\Delta t$ is
preferable although the statistical uncertainty of measuring the joint probabilities
is higher. In practice,
we use a $\Delta t$ that is small enough for a precise finite difference approximation,
and not small enough so that the joint probabilities can be calculated with acceptable accuracy.
In the case of oscillatory EF, the net probability flux is also highly oscillatory,
thus it is convenient to study the integrated probability flux:
\begin{align}
  Q_{J,I} (t) = \int_0^t F_{J,I}(\tau)d \tau,
\end{align}
this expresses the average over time of the effects of the EF on the conformation of the molecule.

@@

\begin{table}
  \centering
  \begin{tabular*}{0.4\textwidth}{@{\extracolsep{\fill}}lc}\hline\hline
    Direction        & time scale [ ps ] \\\hline
    $A_1\rightarrow A_2$        &       $\sim 10$      \\
    $B_1\rightarrow B_2$        &       $\sim 10$      \\    
    $B_2\rightarrow A_1,\ A_2,\ C$        &       $\sim 50$      \\    
    $A_2\rightarrow A_1$        &       $\sim 100$      \\
    $B_2\rightarrow B_1$        &       $\sim 100$      \\    
    $B_1\rightarrow A_1,\ C$        &       $\sim 500$      \\    \hline\hline
  \end{tabular*}
  \caption{A list of main probability fluxes and the corresponding time scales observed in the constant EF case.}
  \label{tab:tmp1}
\end{table}

@@

The Fig.~\ref{fig:tmp3}
shows that the $x$-component of the dipole moment of the alanine dipeptide
reaches 85\% of the maximum value in only 20~ps, which is comparable to
the warm-up time $t_{warm}$, and is 25 times
smaller than the slowest time scale of the conformational relaxation.
Then in the following 400~ps, the dipole slowly relaxes to the 
maximum value, that is, 
6.8~Debye, and  corresponds to a dipole energy of ca. $-13.6$~kJ/mol.
We calculate the averaged dipole moment of different conformations
under the constant EF, and they are 6.8, 6.0, 5.1, 3.1 and 7.1~Debye
for $A_1$, $A_2$, $B_1$, $B_2$ and $C$, respectively.
The dipole energy difference between 3.1 and 7.1~Debye is roughly 8.0~kJ/mol.
The free energy barriers between the
conformations were reported to be of order $20\sim 40$~kJ/mol~\cite{bohner2012algorithm},
which is comparable to the dipole energy difference. 
Therefore, under a constant EF, the system will be driven towards those
conformations with higher dipole moment, because the
energy of the system will be lowered.  This is the reason why
we observe that the $\beta$-sheet conformations with lower dipole moment
are driven towards $\alpha$-helix conformations, 
and $A_1$ and $C$ are conformations with higher probability, because their dipole moments are the highest
among all conformations. Moreover, the $B_2$ conformation vanishes because
its dipole is significantly lower than other conformations.
% This also explains the two-time scale behavior of the molecular dipole
% moment.
From 0 to 20~ps, the probability of
lowest dipole conformation $B_2$ quickly
vanishes. We observe a sharp increment of the average molecular dipole and finally, approaching  500~ps, the conformation slowly migrates from
$B_1$ to $A_1$ and $C$. Moreover, since the dipole moment $A_1$ and $C$
is higher than $B_1$, the average molecular dipole moment slowly increases
to the nonequilibrium steady value.
\subsection{Response of the orientational order parameter}

We also consider the orientation of the alanine dipeptide in space (w.r.t. the direction of EF) as a function
of time.
The geometric direction of the alanine is defined by the red vector in
Fig.~\ref{fig:tmp3}, which is the angle bisector of the two black vectors.
The black vectors are defined in such a way that they connect the $\alpha$-carbon and the carbons
on the methyl groups. We define the orientational order parameter by
\begin{align}
  S_\theta = \langle 3\sin^2\theta - 2\rangle,
\end{align}
where $\theta$ is the angle between the orientation of the alanine dipeptide
and the direction of the EF. The order parameter
indicate how perpendicular does the molecule lie w.r.t. the EF.
If the molecule is perfectly perpendicular to the EF, then $S_\theta = 1$;
If the molecule has no directional
preference , then $S_\theta = 0$;
If the molecule lies perfectly parallel to the EF, then $S_\theta = -2$.
From $t=0$ to roughly 10~ps, the orientational order parameter rapidly
decreases from 0 to $-0.15$, which means a weak alignment of the molecule to the external field.
Then from $t=0$ to 500~ps, the molecule slowly changes to the orientation
that is perpendicular to the EF.
This orientational change is linked to the observed slowest
time scale of the molecular conformational change.

@@

Just as described in Sec.~\ref{sec:tmp2b}, we perform nonequilibrium MD
simulations in a finite size periodic system, and further divided it
into a Hamiltonian dynamics region and a thermostated region.
Since the periodic boundary conditions and the division of the system
are technical (not physical) approximations to the real system with dynamics-preserving thermal control, 
the effect of the finite-size in 
the system settings should be carefully checked.
Therefore, we
perform two additional simulations: one with box dimension $L=4.0$~nm (box size on all directions are the same) and a
Hamiltonian dynamics region of radius $R_{ex} = 1.0$~nm. A second simulation is done with box size $L=4.0$~nm and a Hamiltonian dynamics region of
radius $R_{ex} = 1.5$~nm. We compare the results of these two simulations to those of the system we have used so far, i.e. $L=2.7$~nm and $R_{ex} = 1.0$~nm.
Fig.~\ref{fig:tmp7} presents a satisfactory consistency in
the simulation results of the three systems, which implies that the finite-size effects are
actually negligible.
